---
title: "Samantha Sprague -- HW3 -- MSDS6306"
date: "September 13th, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
sessionInfo()
```


#1. GitHub Cloning (20 points): Using Git, clone the following GitHub repository to your local machine: https://github.com/caesar0301/awesome-public-datasets. In RMarkdown, please show the code (commented out, as it’s not R syntax) that you used to create a new directory, navigate to the appropriate directory, and clone the repository to it. One Git command per line, please.
```{r}
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git init
#Reinitialized existing Git repository in /Users/samanthasprague/testrepo/.git/
#Samanthas-MacBook-Pro:testrepo samanthasprague$ ls
#README.md		SMU_MSDS_Homework
#Samanthas-MacBook-Pro:testrepo samanthasprague$ cd SMU_MSDS_Homework/
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ ls
#DataScienceProfile.png
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ mkdir MSDS6306
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ ls
#DataScienceProfile.png	MSDS6306
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ cd MSDS 6306
#-bash: cd: MSDS: No such file or directory
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ cd MSDS6306
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ ls 
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ mkdir data
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ mkdir hw
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ ls
#data	hw
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ cd data
#Samanthas-MacBook-Pro:data samanthasprague$ git clone https://github.com/caesar0301/awesome-public-datasets
#Cloning into 'awesome-public-datasets'...
#remote: Counting objects: 1484, done.
#remote: Total 1484 (delta 0), reused 0 (delta 0), pack-reused 1484
#Receiving objects: 100% (1484/1484), 565.76 KiB | 1.02 MiB/s, done.
#Resolving deltas: 100% (866/866), done.
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#awesome-public-datasets
```

#Data Summary (20 points): From this aforementioned cloned repo, please extract titanic.csv.zip. To be clear, this does not have to be done in Git or command line.
##a. In R, please read in titanic.csv via either read.table() or read.csv(), assigning it to df. This dataset follows the passengers aboard the Titanic, including their fees paid, rooms rented, and survivorship status.

```{r}
setwd("/Users/samanthasprague/Desktop")
getwd()
titanic<-read.csv("titanic.csv", header=T, sep=",")
head(titanic)
tail(titanic)
str(titanic)
summary(titanic)
```
##b. Output the respective count of females and males aboard the Titanic. Plot the frequency of females and males. Be sure to give an accurate title and label the axes.
```{r}
sex<-titanic$Sex
head(sex)
library(plyr)
count(sex)
plot(sex, ylab="Count of Sex", main="Count of Males and Females aboard the Titanic")
```

##c. Please use one apply function (to review: swirl() modules 11, 12) to output the means of Age, Fare, and Survival. Make sure the output is a real number for all three means.
```{r}
#Null Value Check
colSums(is.na(titanic))
#There are a lot of missing age values, let's remove them from the data set in order to 
#Empty Spaces Check
colSums(titanic=="")
#Unique Values
head(titanic)
age<-titanic$Age
fare<-titanic$Fare
surv<-titanic$Survived
head(age)
head(fare)
head(surv)
length(age)
length(fare)
length(surv)
afs<-data.frame(titanic$Age, titanic$Fare, titanic$Survived)
class(afs)
head(afs)
dim(afs)
sapply(afs, mean)
#Because we get an NA for Age, we need to remove the columns from there dataset where age=NA
afs2<-na.omit(afs)
dim(afs2)
#Just to double check our math, we should be expecting to 177 columns from the original df
891-714
#Math checks out, now time to apply mean to all fields
sapply(afs2, mean)
```

#3. Function Building (30 points): You research sleep and just got your first data set. Later, you’ll have another dataset with the same column names, so you want to create a helper function that you can analyze this dataset and the next. Load sleep_data_01.csv (found at http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/index.html). Questions 3A through 3D should be answered in function(x){}. 3E can be outside of the function.
##a. Create objects for the median Age, the minimum and maximum Duration of sleep, and the mean and standard deviation of the Rosenberg Self Esteem scale (RSES). You may need to specify a few options like in Problem 2 and live session.
```{r}
sleep<-read.csv("sleep_data_01.csv", header=T, sep=",")
head(sleep)
str(sleep)
summary(sleep)
dim(sleep)
colSums(is.na(sleep))
sleep2<-na.omit(sleep)
dim(sleep2)
summary(sleep2)

#Median Function
function_median_age <- function(age) {
  median<- median(age)
  return(median)
}
function_median_age(sleep2$Age)

#Range Function
function_range <- function(duration) {
  range <- range(duration)
  return(range)
}

function_range(sleep2$Duration)

#Mean Function
function_mean <- function(rses){
  mean<-mean(rses)
  return(mean)
}

function_mean(sleep2$RSES)

#SD Function
function_sd <- function(rses) {
  sd<-sd(rses)
  return(sd)
}

function_sd(sleep2$RSES)

```

##b. Create a data.frame object called report: it should consist of the median age, the RSES mean and standard deviation respectively divided by five (since there are five questions and these scores are summed), and the range of Duration (the statistical definition of range; it should be a single number.)
```{r}

function_mean_5 <- function(rses){
  value<-function_mean(rses)/5
  return(value)
}

function_mean_5(sleep2$RSES)

function_sd_5 <- function(rses){
  value <- function_sd(rses)/5
  return(value)
}
function_sd_5(sleep2$RSES)


function_range_sub<- function(duration){
  value<-function_range(duration)
  value2<-value[2]-value[1]
  return(value2)
}
function_range_sub(sleep2$Duration)

report<-data.frame(function_median_age(sleep2$Age), function_mean_5(sleep2$RSES), function_sd_5(sleep2$RSES), function_range_sub(sleep2$Duration))

head(report)

```

##c. Change the column names of this data.frame to MedianAge, SelfEsteem, SE_SD, and DurationRange.
```{r}
names(report) <- c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange")
head(report)
```

#4. FiveThirtyEight Data (30 points): Navigate on GitHub to https://github.com/rudeboybert/fivethirtyeight and read README.md. It will include everything you need.
##a. Install the fivethirtyeight package.
```{r}
#install.packages("fivethirtyeight")
```

##b. In the listing of Data sets in package ‘fivethirtyeight,’ assign the 22nd data set to an object ‘df.’
```{r}
library(fivethirtyeight)
```


```{r}
load("/Users/samanthasprague/Downloads/college_recent_grads.rda")
head(college_recent_grads)
```
```{r}
df<-data.frame(college_recent_grads)
head(df)
```


##c. Use a more detailed list of the data sets to write out the URL in a comment to the related news story.
```{r}
#vignette("fivethirtyeight", package ="fivethirtyeight")
#http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```

##d. Using R command(s), give the dimensions and column names of this data frame.
```{r}
dim(df)
names(df)
str(df)
```

#5. Data Summary (30 points): Use your newly assigned data frame from question 4 for this question.
##a. Write an R command that gives you the column names of the data frame. Right after that, write one that counts the number of columns but not rows. Hint: The number should match one of your numbers in Question 1d for dimensions.
```{r}
#Names of Columns
names(df)
#Number of Columns
ncol(df)
```

##b. Generate a count of each unique major_category in the data frame. I recommend using libraries to help. To be clear, this should look like a matrix or data frame containing the major_category and the frequency it occurs in the dataset. Assign it to major_count.
```{r}
library(stringr)
x<-unique(df$major_category)
length(unique(df$major_category))
y<-str_count(unique(df$major_category))
y
x
major_count_df<-data.frame(x,y)
major_count_df
class(major_count_df)
```

##c. To make things easier to read, put par(las=2) before your plot to make the text perpendicular to the axis. Make a barplot of major_count. Make sure to label the title with something informative (check the vignette if you need), label the x and y axis, and make it any color other than grey. Assign the major_category labels to their respective bar. Flip the barplot horizontally so that bars extend to the right, not upward. All of these options can be done in a single pass of barplot(). Note: It’s okay if it’s wider than the preview pane.
```{r}
?barplot
plot<-barplot(major_count_df$y, col='red', xlab="Majors", ylab="Count of Majors", main="Count of Majors", horiz = TRUE)
?mtext
mtext(major_count_df$x, side=2, at = major_count_df$y, font =1, adj=1, las=2, line=.5, cex.lab=.1)
```

##d. Write the fivethirtyeight data to a csv file. Make sure that it does not have row labels.
```{r}
write.csv(df, file = "hw3.csv", row.names=FALSE)
```

#6. Codebook (30 points):
##a. Start a new repository on GitHub for your SMU MSDS homework. On your local device, make sure there is a directory for Homework at the minimum; you are welcome to add whatever you would like to this repo in addition to your requirements here.
```{r}
#Samanthas-MacBook-Pro:testrepo samanthasprague$ mkdir SMU_MSDS_Homework
#Samanthas-MacBook-Pro:testrepo samanthasprague$ ls
#README.md		SMU_MSDS_Homework
#Samanthas-MacBook-Pro:testrepo samanthasprague$ ls
#README.md		SMU_MSDS_Homework
#Samanthas-MacBook-Pro:testrepo samanthasprague$ cd SMU_MSDS_Homework/
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ ls
#DataScienceProfile.png	MSDS6306
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ cd MSDS6306/
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ ls
#data	hw
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ cd hw
#Samanthas-MacBook-Pro:hw samanthasprague$ echo This is where my R Markdown files will be stored for class MSDS6306. > READMEHW.md
#Samanthas-MacBook-Pro:hw samanthasprague$ ls
#READMEHW.md
#Samanthas-MacBook-Pro:hw samanthasprague$ cd .. 
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ ls
#data	hw
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ cd data
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#awesome-public-datasets
#Samanthas-MacBook-Pro:data samanthasprague$ echo This is where my datasets used in my R #Markdown HW files will be stored. > READMEDATA.md
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#READMEDATA.md		awesome-public-datasets
#Samanthas-MacBook-Pro:hw samanthasprague$ cd ..
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ cd ..
#Samanthas-MacBook-Pro:SMU_MSDS_Homework samanthasprague$ cd ..
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git init
#Reinitialized existing Git repository in /Users/samanthasprague/testrepo/.git/
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git add .
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git commit -m "added readme's"
#[master 98c416c] added readme's
# Committer: Samantha Sprague <samanthasprague@Samanthas-MacBook-Pro.local>
#Your name and email address were configured automatically based
#on your username and hostname. Please check that they are accurate.
#You can suppress this message by setting them explicitly. Run the
#following command and follow the instructions in your editor to edit
#your configuration file:

#    git config --global --edit

#After doing this, you may fix the identity used for this commit with:

#    git commit --amend --reset-author

# 2 files changed, 2 insertions(+)
# create mode 100644 SMU_MSDS_Homework/MSDS6306/data/READMEDATA.md
# create mode 100644 SMU_MSDS_Homework/MSDS6306/hw/READMEHW.md
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git push -u origin master
#Counting objects: 8, done.
#Delta compression using up to 8 threads.
#Compressing objects: 100% (7/7), done.
#Writing objects: 100% (8/8), 763 bytes | 763.00 KiB/s, done.
#Total 8 (delta 0), reused 0 (delta 0)
#To https://github.com/samanthasprague/testrepo.git
#   8cc1c2e..98c416c  master -> master
#Branch master set up to track remote branch master from origin.
```

##b. Create a README.md file which explains the purpose of the repository, the topics included, the sources for the material you post, and contact information in case of questions. Remember, the one in the root directory should be general. You are welcome to make short READMEs for each assignment individually in other folders.
Done. 

##c. In one (or more) of the nested directories, post your RMarkdown script, HTML file, and data from ‘fivethirtyeight.’ Make sure that in your README or elsewhere that you credit fivethirtyeight in some way.
```{r}
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ cd hw
#Samanthas-MacBook-Pro:hw samanthasprague$ ls
#READMEHW.md
#Samanthas-MacBook-Pro:hw samanthasprague$ mkdir hw3
#Samanthas-MacBook-Pro:hw samanthasprague$ ls
#READMEHW.md	hw3
#Samanthas-MacBook-Pro:hw samanthasprague$ cd hw3
#Samanthas-MacBook-Pro:hw3 samanthasprague$ ls
#Samanthas-MacBook-Pro:hw3 samanthasprague$ echo This is the 3rd assignment for MSDS6306. I want to give credit to fivethirtyeight for their datasets and helpful readme files. You can find instructions on how to reach them via the markdown file in this folder, or navigate to the data folder. > READMEhw3.md
#Samanthas-MacBook-Pro:hw3 samanthasprague$ ls
#READMEhw3.md
#Samanthas-MacBook-Pro:hw3 samanthasprague$ cd .. 
#Samanthas-MacBook-Pro:hw samanthasprague$ cd ..
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ ls
#data	hw
#Samanthas-MacBook-Pro:MSDS6306 samanthasprague$ cd data
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#READMEDATA.md		awesome-public-datasets
#Samanthas-MacBook-Pro:data samanthasprague$ mv #/Users/samanthasprague/Desktop/sleep_data_01.csv #/Users/samanthasprague/testrepo/SMU_MSDS_Homework/MSDS6306/data/
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#READMEDATA.md		awesome-public-datasets	sleep_data_01.csv
#Samanthas-MacBook-Pro:data samanthasprague$ mkdir hw3data
#Samanthas-MacBook-Pro:data samanthasprague$ mv sleep_data_01.csv #/Users/samanthasprague/testrepo/SMU_MSDS_Homework/MSDS6306/data/hw3data/
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#READMEDATA.md		awesome-public-datasets	hw3data
#Samanthas-MacBook-Pro:data samanthasprague$ mv awesome-public-datasets/ #/Users/samanthasprague/testrepo/SMU_MSDS_Homework/MSDS6306/data/hw3data/
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#READMEDATA.md	hw3data
#Samanthas-MacBook-Pro:data samanthasprague$ cd hw2
#-bash: cd: hw2: No such file or directory
#Samanthas-MacBook-Pro:data samanthasprague$ cd hw3
#-bash: cd: hw3: No such file or directory
#Samanthas-MacBook-Pro:data samanthasprague$ ls
#READMEDATA.md	hw3data
#Samanthas-MacBook-Pro:data samanthasprague$ cd hw3data/
#Samanthas-MacBook-Pro:hw3data samanthasprague$ ls
#awesome-public-datasets	sleep_data_01.csv
#Samanthas-MacBook-Pro:hw3data samanthasprague$ mv /Users/samanthasprague/Downloads/college_recent_grads.rda /Users/samanthasprague/testrepo/SMU_MSDS_Homework/MSDS6306/data/hw3data/
#Samanthas-MacBook-Pro:hw3data samanthasprague$ ls
#awesome-public-datasets		sleep_data_01.csv
#college_recent_grads.rda
#Samanthas-MacBook-Pro:hw3data samanthasprague$ cd ..
#Samanthas-MacBook-Pro:data samanthasprague$ cd ../../..
#Samanthas-MacBook-Pro:testrepo samanthasprague$ ls
#README.md		SMU_MSDS_Homework
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git add .
#warning: adding embedded git repository: SMU_MSDS_Homework/MSDS6306/data/hw3data/awesome-public-datasets
#hint: You've added another git repository inside your current repository.
#hint: Clones of the outer repository will not contain the contents of
#hint: the embedded repository and will not know how to obtain it.
#hint: If you meant to add a submodule, use:
#hint: 
#hint: 	git submodule add <url> SMU_MSDS_Homework/MSDS6306/data/hw3data/awesome-public-datasets
#hint: 
#hint: If you added this path by mistake, you can remove it from the
#hint: index with:
#hint: 
#hint: 	git rm --cached SMU_MSDS_Homework/MSDS6306/data/hw3data/awesome-public-datasets
#hint: 
#hint: See "git help submodule" for more information.
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git commit -m "added datasets"
#[master e46440d] added datasets
# Committer: Samantha Sprague <samanthasprague@Samanthas-MacBook-Pro.local>
#Your name and email address were configured automatically based
#on your username and hostname. Please check that they are accurate.
#You can suppress this message by setting them explicitly. Run the
#following command and follow the instructions in your editor to edit
#your configuration file:

#    git config --global --edit

#After doing this, you may fix the identity used for this commit with:

#    git commit --amend --reset-author

# 4 files changed, 103 insertions(+)
# rename SMU_MSDS_Homework/MSDS6306/data/{ => hw3data}/awesome-public-datasets (100%)
# create mode 100644 SMU_MSDS_Homework/MSDS6306/data/hw3data/college_recent_grads.rda
# create mode 100644 SMU_MSDS_Homework/MSDS6306/data/hw3data/sleep_data_01.csv
# create mode 100644 SMU_MSDS_Homework/MSDS6306/hw/hw3/READMEhw3.md
#Samanthas-MacBook-Pro:testrepo samanthasprague$ git push -u origin master
#Counting objects: 11, done.
#Delta compression using up to 8 threads.
#Compressing objects: 100% (10/10), done.
#Writing objects: 100% (11/11), 14.67 KiB | 7.34 MiB/s, done.
#Total 11 (delta 0), reused 0 (delta 0)
#To https://github.com/samanthasprague/testrepo.git
#   fd1c357..e46440d  master -> master
#Branch master set up to track remote branch master from origin.
#Samanthas-MacBook-Pro:testrepo samanthasprague$ 
```


##d. In your RMarkdown script, please provide the link to this GitHub so the grader can see it.
```{r}
#https://github.com/samanthasprague/testrepo
```

